Implicit yaddayadda

### 2011-09-23 papers/PLDI2012/implicitization.tex
exemple pour la technique combinée

\usepackage{amsmath,amsfonts}
\usepackage{dmnatbib}
\usepackage[mathscr]{euscript}
\usepackage{algorithmic,algorithm}

@@

\newcommand{\abstr}[1]{#1^\sharp}
\newcommand{\parts}[1]{\mathscr{P}(#1)}
\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\QQ}{\mathbb{Q}}

@@

\subsection{A simple, motivating example}
Consider the following program, adapted from \cite{Monniaux_Gonnord_SAS11}:
\lstinputlisting{rlim_int.c}
This program implements a construct commonly found in control programs (in e.g. automotive or avionics): a rate limiter. For the sake of simplicity, we chose it to be fed a nondeterministic input clamped between $[-100000,100000]$, but in a real system it would be integrated in a reactive control loop and its input connected to a complex system with unknown output range.

@@

@@

Our improvement consists in a slight modification of the SMT formula we solve in
\ref{sec:guided_multigraph}.
We introduce in this formula Boolean variables $\{d_k, 1 \leq k \leq m\}$, so
that we can easily find in the model which abstract value of the disjunction we
have to choose. The resulting formula that is given to the SMT solver is defined
by $g(p_i)$.

@@

Our algorithm for computing disjunctive invariants is described in Algorithm
\ref{algo:disjunctive}

\begin{algorithm}[!h]
\caption{Disjunctive invariant computation with implicit paths}\label{gulwani2}
\label{algo:disjunctive}
\begin{algorithmic}[1] 
\FORALL {$k \in \{1,..,m\} \setminus \{\delta\}$}
	\STATE $X_{i,k} \gets \perp$
\ENDFOR
\STATE $X_{i,\delta} \gets I_{p_i}$
\WHILE {true}
		\STATE $res \gets SmtSolve\left[g(p_i)\right]$
	\IF {$res = unsat$}
		\STATE \textbf{break}
	\ENDIF
	\STATE Compute the path $\tau_{i,k}$ from $res$ 
	\STATE Take $j \in \{ l | d_l = true\}$ 
	\STATE $X_{i,\sigma(j,k)} \gets X_{i,\sigma(j,k)} \sqcup
	\tau_{i,k}(X_{i,j})$
\ENDWHILE
\end{algorithmic}
\end{algorithm}

>>>>

\usepackage{amsmath,amsfonts}
\usepackage{mathcomp,amssymb}
\usepackage{dmnatbib}
\usepackage[mathscr]{euscript}
\usepackage{algorithmic,algorithm}

@@

\newcommand{\abstr}[1]{#1^\sharp}
\newcommand{\parts}[1]{\mathscr{P}(#1)}
\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\QQ}{\mathbb{Q}}
\newcommand{\widening}{\mathop{\triangledown}}

@@

\subsection{A simple, motivating example}
\label{subsec:rate_lim}
Consider the following program, adapted from \cite{Monniaux_Gonnord_SAS11}:
\lstinputlisting{rlim_int.c}
This program implements a construct commonly found in control programs (in e.g. automotive or avionics): a rate limiter. For the sake of simplicity, we chose it to be fed a nondeterministic input clamped between $[-100000,100000]$, but in a real system it would be integrated in a reactive control loop and its input connected to a complex system with unknown output range.

@@

\subsection{Example}

We revise the rate limiter described in \ref{subsec:rate_lim}. In this example,
\emph{Path-focused} technique works well because all the paths starting at the
loop header are actually self loops. In such a case, the technique performs a
widening/narrowing sequence or accelerates the loop, thus leading to a precise
invariant. However, in real cases, there also exist paths that are not
self loops, and \emph{Path-focusing} only performs a widening. 
This widening may induce unrecoverable loss of precision.

Suppose the main loop of the rate limiter contains a nested loop:
\lstinputlisting{rlim_int_loop.c}

We choose $P_R$ as the set of loop headers of the function, plus the initial
state. In this case, we have three elements in $P_R$.
We also unroll the loop once, in order to distinguish the paths that don't go
through the loop from the others:
\begin{lstlisting}[numbers=none]
if (wait) {
   while (wait()) {}
}
\end{lstlisting}

The main loop in the expanded multigraph has then 24 distinct paths, half of
them being self loops, and the other half going to the header of the nested loop.

Guided static analysis from \citet{DBLP:conf/sas/GopanR07} yields at line 5 
$\lstinline|x_old| \in (-\infty,+\infty)$.
Path-focused technique from \citet{Monniaux_Gonnord_SAS11} performs slightly
better, and finds $\lstinline|x_old| \in (-\infty,10000)$.
Our combined technique gives us the expected invariant 
$\lstinline|x_old| \in (-10000,10000)$.

@@

Our improvement consists in a slight modification of the SMT formula we solve in
\ref{sec:guided_multigraph}.
We introduce in this formula Boolean variables $\{d_j, 1 \leq j \leq m\}$, so
that we can easily find in the model which abstract value of the disjunction we
have to choose. The resulting formula that is given to the SMT solver is defined
by $g(p_i)$.
When the formula is satisfiable, we know that the index $j$ of the starting
disjunct that has to be chosen is the one for which the associate Boolean value
$d_j$ is \emph{true} in the model. Then, we can easily compute the value of 
$\sigma(j,k)$, thus know the index of the disjunct to join with.

@@

Our algorithm for computing disjunctive invariants is described in Algorithm
\ref{algo:disjunctive}. In this algorithm, instead of simply joining the two
abstract values, we use the widening operator to ensure the algorithm terminates
when using abstract domains with infinite height.

\begin{algorithm}[!h]
\caption{Disjunctive invariant computation with implicit paths}\label{gulwani2}
\label{algo:disjunctive}
\begin{algorithmic}[1] 
\FORALL {$k \in \{1,..,m\} \setminus \{\delta\}$}
	\STATE $X_{i,k} \gets \perp$
\ENDFOR
\STATE $X_{i,\delta} \gets I_{p_i}$
\WHILE {true}
		\STATE $res \gets SmtSolve\left[g(p_i)\right]$
	\IF {$res = unsat$}
		\STATE \textbf{break}
	\ENDIF
	\STATE Compute the path $\tau_{i,k}$ from $res$ 
	\STATE Take $j \in \{ l | d_l = true\}$ 
	\STATE $X_{i,\sigma(j,k)} \gets X_{i,\sigma(j,k)} \widening (X_{i,\sigma(j,k)} \sqcup \tau_{i,k}(X_{i,j}))$
\ENDWHILE
\end{algorithmic}
\end{algorithm}
